---
title: |
 | Enhance and Advance
 | Introduction to R
subtitle: "Data Visualization Application: Gapminder" 
author: Eric Dunford
affiliation: "Georgetown University"
date: "Summer 2020" 
output: 
  html_notebook:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include= F}
knitr::opts_chunk$set(error=F,warning = F,comment=F)

# Load dependencies--------------------------

# The tidyverse package covered last time 

#install.packages("tidyverse")
require(tidyverse)

# install.packages("ggthemes")
require(ggthemes) # for great visualization colors and themes. 

# install.packages("maps")
require(maps) # for some maps data

# Gapminder data (for example)
#install.packages("gapminder")
require(gapminder)
```

# Data 

The [`Gapminder` dataset](https://www.gapminder.org/) is a famous dataset used by Hans Rosling to visualize development outcomes. The data covers 1952 to 2007 in five year intervals and measures life expectancy, population, and GDP Per Capita. 

```{r}
gapminder %>% head()
```


# Get to know the data through visualization

Each of the following questions are targeted at making sure we understand our data better. A great way to get a "feel" for a dataset is to visualize it. Answer each of the below questions with a (publishable-quality) picture. 


#### 1. How is `lifeExp`, `pop` and `gdpPercap` variables distributed?


#### 2. What's the relationship between economic development and life expectancy? Is the relationship the same for all continents?


#### 3. Which countries in Africa have the lowest levels of life expectancy?





