---
title: |
 | Enhance and Advance
 | Introduction to R
subtitle: "Data Visualization Application: Gapminder" 
author: Eric Dunford
affiliation: "Georgetown University"
date: "Summer 2020" 
output: 
  html_notebook:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include= F}
knitr::opts_chunk$set(error=F,warning = F,comment=F)

# Load dependencies--------------------------

# The tidyverse package covered last time 

#install.packages("tidyverse")
require(tidyverse)

# install.packages("ggthemes")
require(ggthemes) # for great visualization colors and themes. 

# install.packages("maps")
require(maps) # for some maps data

# Gapminder data (for example)
#install.packages("gapminder")
require(gapminder)
```

# Data 

The [`Gapminder` dataset](https://www.gapminder.org/) is a famous dataset used by Hans Rosling to visualize development outcomes. The data covers 1952 to 2007 in five year intervals and measures life expectancy, population, and GDP Per Capita. 

```{r}
gapminder %>% head()
```


# Get to know the data through visualization

Each of the following questions are targeted at making sure we understand our data better. A great way to get a "feel" for a dataset is to visualize it. Answer each of the below questions with a (publishable-quality) picture. 


#### 1. How is `lifeExp`, `pop` and `gdpPercap` variables distributed?

```{r}
gapminder %>% 
  ggplot(aes(x=lifeExp)) +
  geom_histogram(bins=50)


gapminder %>% 
  mutate(ln_pop = log(pop)) %>%
  ggplot(aes(x=ln_pop)) +
  geom_histogram(bins=50)


gapminder %>% 
  mutate(ln_gdppp = log(gdpPercap)) %>%
  ggplot(aes(x=ln_gdppp)) +
  geom_histogram(bins=50)
```


```{r,fig.align="center",fig.height=4,fig.width=10}
# Data manipulation
gapminder %>% 
  mutate(ln_gdppp = log(gdpPercap),
         ln_pop = log(pop)) %>%
  pivot_longer(cols=c(lifeExp,ln_pop,ln_gdppp)) %>% 
  mutate(name = case_when(
    name == "lifeExp" ~ "Life Expectancy",
    name == "ln_pop" ~ "Log Population",
    name == "ln_gdppp" ~ "Log GDP Per Capita"
  )) %>% 
  
  # Plot code
  ggplot(aes(x=value,fill=name)) +
  geom_histogram(bins=25,show.legend = F,alpha=.8,color="white") +
  # geom_density(show.legend = F,alpha=.8,color="white") +
  facet_wrap(~name,scales = "free_x") + 
  scale_fill_economist() +
  labs(caption = "Source: gapminder.org",
       title = "Distribution of variables") +
  theme_fivethirtyeight() +
  theme(text = element_text(family="serif",size=14,face="bold"))
```


#### 2. What's the relationship between economic development and life expectancy? Is the relationship the same for all continents?


```{r,fig.align="center",fig.height=4,fig.width=7}
gapminder %>% 
  mutate(ln_gdppc = log(gdpPercap)) %>% 
  ggplot(aes(x=ln_gdppc,y=lifeExp)) +
  geom_point(alpha=.4,color="grey50") +
  geom_smooth(method = "loess",se=F,color="darkred",size=1.5) +
  theme_minimal()
```


```{r,fig.align="center",fig.height=4,fig.width=7}
gapminder %>% 
  mutate(ln_gdppc = log(gdpPercap)) %>% 
  ggplot(aes(x=ln_gdppc,y=lifeExp,color=continent)) +
  geom_point(alpha=.4) +
  geom_smooth(method = "loess",se=F,size=1.5) +
  theme_minimal() +
  facet_wrap(~continent,nrow=1) +
  theme(legend.position = "bottom")
```


```{r,fig.align="center",fig.height=4,fig.width=7}
gapminder %>% 
  mutate(ln_gdppc = log(gdpPercap)) %>% 
  ggplot(aes(x=ln_gdppc,y=lifeExp,color=continent)) +
  # geom_point(alpha=.4) +
  geom_smooth(method = "loess",se=F,size=1.5) +
  scale_color_gdocs() +
  theme_minimal() +
  labs(color="",x= "Log GDP Per Capita",y="Life Expectancy") +
  theme(legend.position = "bottom")
```

#### 3. Which countries in Africa have the lowest levels of life expectancy?

```{r,fig.align="center",fig.height=7,fig.width=7}
# Data 
gapminder %>% 
  filter(continent == "Africa") %>% 
  group_by(country) %>% 
  summarize(lifeExp = mean(lifeExp),.groups="drop") %>% 
  
  # Plot
  ggplot(aes(x=lifeExp,y=fct_reorder(country,desc(lifeExp)),
             fill = lifeExp) ) +
  geom_col() +
  scale_fill_gradient2_tableau() +
  labs(y="",x="Life Expectancy") +
  theme_hc()
```

```{r,fig.align="center",fig.height=7,fig.width=10}
map_data("world") %>% 
  ggplot(aes(x = long,y=lat,group=group)) +
  geom_polygon()
```




